---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%",
  fig.width = 8,
  fig.height = 6
)

library(png)
library(ggplot2)
library(dplyr)
library(fastpng)


if (FALSE) {
  covr::report(covr::package_coverage(
    line_exclusions = list('src/spng.c', 'src/spng.h')
  ))
}

if (FALSE) {
  # Check for functions which do not have an @examples block for roxygen
  system("grep -c examples man/*Rd", intern = TRUE) |> 
    grep(":0$", x = _, value = TRUE)
  
  checkhelper::check_as_cran()
  checkhelper::find_missing_tags()
}


if (FALSE) {
  covr::report()
}


# Makevars options to do some deep testing for CRAN

# Type conversions are sane
# PKG_FLAG=-Wconversion

# Pointer overflow checks i.e. dodgy pointer arithmetic
# PKG_CFLAGS+=-fsanitize=pointer-overflow -fsanitize-trap=pointer-overflow
# Then run in the debugger:
# R -d lldb 
# run
# testthat::test_local()
```

# fastpng


<!-- badges: start -->
![](https://img.shields.io/badge/cool-useless-green.svg)
[![R-CMD-check](https://github.com/coolbutuseless/fastpng/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/coolbutuseless/fastpng/actions/workflows/R-CMD-check.yaml)
[![CRAN status](https://www.r-pkg.org/badges/version/fastpng)](https://CRAN.R-project.org/package=fastpng)
<!-- badges: end -->

`{fastpng}` reads and writes PNG images.

`{fastpng}` exposes configuration options so that the user can make a trade-off
between speed of writing and PNG size.  These options include:

* Compression level
* Filter use
* Image transposition

For example, writing uncompressed PNG images can be 100x faster than writing with 
regular compression settings.

`fastpng` is an R wrapper for [libspng](https://github.com/randy408/libspng) - current v0.7.4


## Features 

### Supported image data in R

Supported images each have examples in the `test_image` as part of this package.

* **Native rasters**
* **Rasters**
    * With hex colour formats: #RGB, #RGBA, #RRGGBB, #RRGGBBAA
    * Standard R colour names also supported e.g. 'red', 'white'
* **Numeric arrays**
    * Values in range [0,1]
    * 1-, 2-, 3- and 4-plane numeric arrays (interpreted as gray, gray+alpha,
      RGB and RGBA images)
* **Integer arrays**
    * Values in range [0,255] treated as 8-bit values
    * Values in range [0,65535] treated as 16-bit for PNG writing
* **Integer matrix + an indexed palette of colours** 
* **Raw vectors** with a specification for data layout


### Supported PNG image types

* 8-bit and 16-bit PNGs
* RGBA, RGB, Gray + Alpha, Gray PNGs
* Indexed colour PNGs
* RGB PNGs with a specified transparency colour (using [tRNS chunk](https://www.w3.org/TR/PNG-Chunks))

### Comparison to standard `{png}` library

|             |  `{fastpng}`    |     `{png}`   |
|:------------|-----------------|---------------|
| Numeric arrays              |    Yes        |    Yes       |
| Native rasters              |    Yes        |    Yes       |
| Rasters                     |    Yes        |              |
| Integer Arrays              |    Yes        |              |
| Indexed PNGs                |    Yes        |              |
| `tRNS` transparency         |    Yes        |              |
| Configurable compression    |    Yes        |              |
| Configurable filtering      |    Yes        |              |
| Configurable transposition  |    Yes        |              |


## Compression Settings: Speed / size tradeoff


<details>
<summary> Click to reveal benchmark code and results table </summary>

```{r echo=TRUE}
library(png)
im <- test_image$array$rgba

res <- bench::mark(
  fastpng::write_png(im, compression_level =  0),
  fastpng::write_png(im, compression_level =  1),
  fastpng::write_png(im, compression_level =  2),
  fastpng::write_png(im, compression_level =  3),
  fastpng::write_png(im, compression_level =  4),
  fastpng::write_png(im, compression_level =  5),
  fastpng::write_png(im, compression_level =  6),
  fastpng::write_png(im, compression_level =  7),
  fastpng::write_png(im, compression_level =  8),
  fastpng::write_png(im, compression_level =  9),
  check = FALSE,
  relative = FALSE
)

res <- res %>% 
  select(writes_per_second = `itr/sec`) %>%
  mutate(
    compression = 0:9,
    package = "fastpng"
  )


sizes <- vapply(0:9, \(x) fastpng::write_png(im, compression_level = x) |> length(), integer(1))
df <- data.frame(compression = 0:9, size = sizes)


plot_df <- left_join(res, df, by = 'compression')


png_speed <- bench::mark(writePNG(im))$`itr/sec`
png_size  <- length(writePNG(im))

plot_df <- plot_df %>% 
  add_row(writes_per_second = png_speed, size = png_size, package = "png") %>%
  mutate(
    compression_ratio = prod(dim(im)) * 8 / size
  )

knitr::kable(plot_df)
```

</details>

```{r echo=FALSE}
p <- ggplot(plot_df) +
  geom_point(aes(compression_ratio, writes_per_second, colour = package)) + 
  geom_text(aes(compression_ratio, writes_per_second, label = compression), na.rm = TRUE, hjust = -0.5, vjust = -0.5) +
  theme_bw() + 
  scale_y_log10(breaks = c(1, 5, 10, 50, 100, 500, 1000, 5000)) +
  labs(
    x = "Compression Ratio (Higher values are better)",
    y = "Writes per second (Bigger is better)",
    title = "Speed-Compression tradeoff writing PNGs from array data",
    subtitle = paste(
      "Numbers indicate 'compression_level' setting in {fastpng}",
      "Lower `compression_level' settings mean faster writes", 
      "With compression disabled, {fastpng} can write test image ~6000 times per second",
      sep = "\n"
    )
  ) +
  theme(
    legend.position = 'bottom'
  )


ggsave("vignettes/image/benchmark.png")
p
```


## Installation

This package can be installed from CRAN

``` r
install.packages('fastpng')
```

You can install the latest development version from
[GitHub](https://github.com/coolbutuseless/fastpng) with:

``` r
# install.package('remotes')
remotes::install_github('coolbutuseless/fastpng')
```

Pre-built source/binary versions can also be installed from
[R-universe](https://r-universe.dev)

``` r
install.packages('fastpng', repos = c('https://coolbutuseless.r-universe.dev', 'https://cloud.r-project.org'))
```

## What's in the box

* `read_png()` to read a PNG from a file or a raw vector
* `write_png()` to write an R image as a PNG file or PNG data in a raw vector
* `get_png_info()` - interrogate a vector of raw values containing a PNG image
  to determine image information i.e. width, height, bit_depth, color_type, 
  compression_method, filter_method, interlace_method.
* `test_image` is a list of images.  These are images contained in different 
  datastructures and of differing bitdepth etc:
  RGBA and RGB numeric arrays, raster, native raster.
  

## Example: Read a PNG into R

```{r fig.height = 2}
library(fastpng)
png_file <- system.file("img", "Rlogo.png", package="png")
fastpng::get_png_info(png_file)

ras <- read_png(png_file, type = 'raster') 
grid::grid.raster(ras, interpolate = FALSE)
```


#### Read as a raster (of hex colours)

```{r}
ras <- read_png(png_file, type = "raster")
ras[7:11, 79:83]
```

#### Read as a numeric array

```{r}
ras <- read_png(png_file, type = "array")
ras[7:11, 79:83, 1] # red channel
```

#### Read as an integer array

```{r}
ras <- read_png(png_file, type = "array", array_type = 'integer')
ras[7:11, 79:83, 1] # red channel
```

#### Read as a native raster

```{r}
im <- read_png(png_file, type = "nativeraster")
im[7:11, 79:83]
```

## Write an image to PNG with/without compression

```{r}
png_file <- tempfile()
write_png(im, png_file)  # standard compression
file.size(png_file)
```

```{r}
write_png(im, png_file, compression_level = 0) # no compression, but fast!
file.size(png_file)
```



## Write integer matrix as indexed PNG

```{r}
indices <- test_image$indexed$integer_index
palette <- test_image$indexed$palette

dim(indices)
indices[1:10, 1:10]
palette[1:10]
```


```{r eval=FALSE}
tmp <- tempfile()
write_png(image = indices, palette = palette, file = tmp)
```

```{r echo=FALSE, fig.height = 2}
write_png(image = indices, palette = palette) |>
  read_png() |> 
  grid::grid.raster()
```



```{r echo=FALSE, eval=FALSE}
write_png(image = indices, palette = palette, file = "working/indexed.png")
```



