---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "60%"
)

library(spng)
```

# spng


<!-- badges: start -->
![](https://img.shields.io/badge/cool-useless-green.svg)
[![R build status](https://github.com/coolbutuseless/spng/workflows/R-CMD-check/badge.svg)](https://github.com/coolbutuseless/spng/actions)
<!-- badges: end -->

`spng` offers in-memory decompression of PNG images to vectors of raw values.

This is useful if you have bytes representing a PNG image (e.g. from a database)
and need to decompress these to an array representation within R.

`spng` is a R wrapper for [libspng](https://github.com/randy408/libspng) - current v0.7.4



* [libspng API docs](https://libspng.org/docs/api/)


## To Do

* Test Image Suite 
    * http://www.schaik.com/pngsuite/
* Read from `file` as well as `raw`
* Implement functions `write_png()` and `read_png()` which 
  will dispatch on R object type.
* `raster` images can be RGB or RGBA.  RGBA rasters are supported.
  Add support for RGB rasters


## Installation

You can install from [GitHub](https://github.com/coolbutuseless/spng) with:

``` r
# install.package('remotes')
remotes::install_github('coolbutuseless/spng')
```

## What's in the box

* `read_png_as_raw(raw_vec, fmt, flags)` 
* `read_png_as_nara(raw_vec, flags)`
* `read_png_as_raster(raw_vec, flags)`
* `read_png_as_rgba(raw_vec, flags)`
* `read_png_as_rgb(raw_vec, flags)`
   
* `extract_png_info(raw_vec)` - interrogate a vector of raw values containing a PNG image
  to determine image information i.e. width, height, bit_depth, color_type, 
  compression_method, filter_method, interlace_method.
  

## Example: Decompress a PNG in memory

```{r example}
library(spng)

#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# A PNG file everyone should have!
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
png_file <- system.file("img", "Rlogo.png", package="png")
img <- magick::image_read(png_file)
img
```


```{r }
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# Read in the raw bytes
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
png_data <- readBin(png_file, 'raw', n = file.size(png_file))
png_data[1:100]
```


```{r}
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# Get info about the PNG 
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
(extract_png_info <- spng::extract_png_info(png_data))
```


### Read PNG as native raster

```{r}
nara <- read_png_as_nara(png_data)
grid::grid.raster(nara, interpolate = FALSE)
```


```{r eval=FALSE, echo=FALSE}
library(png)
bench::mark(
  read_png_as_raw(png_data),
  read_png_as_nara(png_data),
  read_png_as_raster(png_data),
  read_png_as_rgba(png_data),
  read_png_as_rgb(png_data),
  readPNG(png_data, native = FALSE),
  readPNG(png_data, native = FALSE) |> as.raster(),
  readPNG(png_data, native = TRUE),
  check = FALSE
)
```



```{r echo=FALSE, eval = FALSE}
im <- matrix("#445566FF", 100, 150)
as.raster(im)
grid.raster(im)
```


### Read PNG as raster

```{r}
ras <- read_png_as_raster(png_data)
plot(ras, interpolate = FALSE)
```


### Read PNG as RGBA array

```{r}
arr <- read_png_as_rgba(png_data)
plot(as.raster(arr), interpolate = FALSE)
```



### Read PNG as RGB array

Ignoring any alpha channel in the image.

```{r}
arr <- read_png_as_rgb(png_data)
plot(as.raster(arr), interpolate = FALSE)
```



### Read PNG as raw vector

```{r}
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# Unpack the raw PNG bytes into RGBA 8-bits-per-color packed format. 
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
img_data <- spng::read_png_as_raw(png_data, fmt = spng_format$SPNG_FMT_RGBA8)
img_data[1:200]
```




## Acknowledgements

* R Core for developing and maintaining the language.
* CRAN maintainers, for patiently shepherding packages onto CRAN and maintaining
  the repository
