---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%",
  fig.width = 8,
  fig.height = 6
)

library(fastpng)

# Makevars options to do some deep testing for CRAN

# Type conversions are sane
# PKG_FLAG=-Wconversion

# Pointer overflow checks i.e. dodgy pointer arithmetic
# PKG_CFLAGS+=-fsanitize=pointer-overflow -fsanitize-trap=pointer-overflow
# Then run in the debugger:
# R -d lldb 
# run
# testthat::test_local()
```

# fastpng


<!-- badges: start -->
![](https://img.shields.io/badge/cool-useless-green.svg)
[![R-CMD-check](https://github.com/coolbutuseless/fastpng/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/coolbutuseless/fastpng/actions/workflows/R-CMD-check.yaml)
<!-- badges: end -->

`{fastpng}` provides functions for read/write of PNG image from files and raw vectors.

By disabling compression when writing image data to PNG, this can ben ~75x faster
than standard PNG packages.

`{fastpng}` 

* Provides explicit support for images as rasters, native rasters, numeric arrays (RGBA, RGB, and Grayscale)
  and indexed images
* Flags to configure reading of PNG
    * gamma correction
* Flags to configure writing of PNG
    * Compression level
    * PNG filter settings

`fastpng` is an R wrapper for [libspng](https://github.com/randy408/libspng) - current v0.7.4

* [libspng API docs](https://libspng.org/docs/api/)


```{r warning=FALSE, echo=FALSE}
library(ggplot2)

image <- test_image$array$rgba

res <- bench::mark(
  png::writePNG(image),
  fastpng::write_png(image),
  fastpng::write_png(image, compression_level = 0),
  check = FALSE,
  relative = TRUE
)

res$name <- c("{png}", "{fastpng}", "{fastpng}\nNo compression")
res$name <- factor(res$name, levels = res$name)

ggplot(res) + 
  geom_col(aes(name, `itr/sec`)) + 
  theme_bw(14) + 
  labs(
    title = "Relative speed of writing 300x400 RGBA array to PNG",
    subtitle = "Bigger is better",
    y = "Relative speed",
    x = NULL
  )
```





```{r warning=FALSE, echo=FALSE}
library(ggplot2)

test_image_gray8 <- test_image$array$gray

res <- bench::mark(
  png::writePNG(test_image_gray8),
  fastpng::write_png(test_image_gray8),
  fastpng::write_png(test_image_gray8, compression_level = 0),
  fastpng::write_png(test_image_gray8, compression_level = 0, avoid_traanspose = TRUE),
  check = FALSE,
  relative = TRUE
)

res$name <- c("{png}", "{fastpng}", "{fastpng}\nNo compression", "{snpg}\nNo compression.\nNo transpose")
res$name <- factor(res$name, levels = res$name)

ggplot(res) + 
  geom_col(aes(name, `itr/sec`)) + 
  theme_bw(14) + 
  labs(
    title = "Relative speed of writing 300x400 grey matrix to PNG",
    subtitle = "Bigger is better",
    y = "Relative speed",
    x = NULL
  )
```


## ToDo

* Transparency via the `tRNS` chunk for RGBA and RGB format PNGs.  (Already
  implemented for indexed images).
* 16 bit support


## Installation

You can install from [GitHub](https://github.com/coolbutuseless/fastpng) with:

``` r
# install.package('remotes')
remotes::install_github('coolbutuseless/fastpng')
```

## What's in the box

* `read_png()` to read a PNG from a file or a raw vector
* `write_png()` to write data as a PNG file or PNG data in a raw vector
* `get_png_info()` - interrogate a vector of raw values containing a PNG image
  to determine image information i.e. width, height, bit_depth, color_type, 
  compression_method, filter_method, interlace_method.
* `test_image` is a named list of different image representations in R:
  RGBA and RGB numeric arrays, raster, native raster.
  
Supported R image types:

* Native Raster (integer matrix with class 'nativeRaster')
* Raster (character matrix with class 'raster') with hex colour values of the 
  form `#RRGGBBAA` or `#RRGGBB`.  Note: R colour names are not supported here.
* 3D numeric array containing RGBA values in the range [0, 1]
* 3D numeric array containing RGB values in the range [0, 1]
* 2D numeric matrix containing greyscale values in the range [0, 1] 
* Integer 2D matrix paired with a colour palette (specified as a character vector of hex colours).
  Also supports alpha channel.
  

## Example: Decompress a PNG from a raw vector

```{r example}
library(fastpng)

#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# A PNG file everyone should have!
# Read in the raw bytes
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
png_file <- system.file("img", "Rlogo.png", package="png")
png_data <- readBin(png_file, 'raw', n = file.size(png_file))
png_data[1:100]
```


```{r}
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# Get info about the PNG 
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
(get_png_info <- fastpng::get_png_info(png_data))
```


### Read PNG as native raster

```{r}
nara <- read_png(png_data, type = 'native_raster')
nara[1:10, 1:10]
grid::grid.raster(nara, interpolate = FALSE)
```


```{r eval=FALSE, echo=FALSE}
library(png)
bench::mark(
  # read_png(png_data),
  read_png(png_data, type = 'native_raster'),
  read_png(png_data, type = 'ras'),
  read_png(png_data, type = 'rgba'),
  read_png(png_data, type = 'rgb'),
  readPNG(png_data, native = FALSE),
  readPNG(png_data, native = FALSE) |> as.raster(),
  readPNG(png_data, native = TRUE),
  check = FALSE
)
```



```{r echo=FALSE, eval = FALSE}
im <- matrix("#445566FF", 100, 150)
as.raster(im)
grid.raster(im)
```


### Write image as indexed PNG

```{r}
indices <- test_image$indexed$integer_index
palette <- test_image$indexed$palette

dim(indices)
indices[1:10, 1:10]
palette[1:10]
```


```{r echo=FALSE}
tmp <- tempfile()
write_png(image = indices, palette = palette, file = tmp)
```

```{r echo=FALSE}
write_png(image = indices, palette = palette) |>
  read_png() |> 
  grid::grid.raster()
```



```{r echo=FALSE, eval=FALSE}
write_png(image = indices, palette = palette, file = "working/indexed.png")
```






## Acknowledgements

* R Core for developing and maintaining the language.
* CRAN maintainers, for patiently shepherding packages onto CRAN and maintaining
  the repository
