---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)

library(spng)
```

# spng


<!-- badges: start -->
![](https://img.shields.io/badge/cool-useless-green.svg)
<!-- badges: end -->

`spng` is an R wrapper for the [libspng](https://github.com/randy408/libspng).

There's nothing much here yet except for basic decoding of PNG bytes to a raw vector.


* [libpng API docs](https://libspng.org/docs/api/)


## Installation

You can install from [GitHub](https://github.com/coolbutuseless/spng) with:

``` r
# install.package('remotes')
remotes::install_github('coolbutuseless/spng')
```

## Example: Decompress a PNG in memory

```{r example, out.width="60%"}
library(spng)

#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# A PNG file everyone should have!
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
png_file <- system.file("img", "Rlogo.png", package="png")
img <- magick::image_read(png_file)
img
```


```{r out.width="60%"}
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# Read in the raw bytes
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
png_data <- readBin(png_file, 'raw', n = file.size(png_file))
png_data[1:100]
```


```{r}
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# Get info about the PNG 
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
(png_info <- spng::png_info(png_data))
```


```{r out.width="60%"}
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# Unpack the raw PNG bytes into ABGR32 (I think) packed pixel format
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
img_data <- spng::depng(png_data)
img_data[1:200]

```


```{r out.width="60%"}
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# Pick the green channel and plot as greyscale
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
N <- length(img_data)
mat <- matrix(img_data[seq(2, N, 4)], nrow = png_info$width, ncol = png_info$height)
plot(as.raster(mat))
```


```{r out.width="60%"}
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# Construct an array by plucking the relevant bytes.
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
arr <- array(
  c(
    img_data[seq(1, N, 4)], # R
    img_data[seq(2, N, 4)], # G
    img_data[seq(3, N, 4)]  # B
  ),
  dim = c(png_info$width, png_info$height, 3)
  
)

plot(as.raster(arr))
```


## Related Software


## Acknowledgements

* R Core for developing and maintaining the language.
* CRAN maintainers, for patiently shepherding packages onto CRAN and maintaining
  the repository
